Refactoring the Candidate Cancer Gene Database

### 2021-05-21 presentation.Rmd
final edits

> - Cancer bio is complicated.
> - Driver genes present primary targets.
> - Screening methods...

@@

@@

> - Create an interactive web tool to quickly refine the stated data set by a variety of parameters.
> - Draw from universal resources to define homology relationships for candidate genes between animal models.
> - Curate disease classification and meta-analysis through complete CIS listing and relative rank and strength scores.
> - Offer easy access to external tools from filtered results for deeper biological context.

<div class="notes">
- relative rank was based either on the number of insertions in a given CIS or the P-value associated with the given CIS
- The relative rank was calculated by ranking all CISs in a given study and then assigning grades
- We assigned a letter grade of A for the top 10%, B for the top 11–25%, C for the top 26–50% and D for the bottom 50%.
</div>

@@

<div class="col">
> - Filtered results are downloadable.
> - Wrapped in a simple, instructional website built with classic LAMP stack.
> - Hosted by UMN-OIT.
</div>

@@

> - Machine displacement in provisioning by UMN-OIT
> - Use of abandonded, proprietary widgets
> - Outsized SQL database

@@

- <span style="color: #a3be8c;">**Machine displacement in provisioning by UMN-OIT**</span>
- Use of abandonded, proprietary widgets
- <span style="color: #a3be8c;">**Outsized SQL database**</span>

@@

<div class="notes">
- ADD SOME NOTES
</div>

@@

@@

<div class="notes">
- codebase simplified in a few ways.
- build table process was dramatically streamlined.
</div>

```{r echo = T}
export <- homogTable %>%
  full_join(df) %>%
  group_by(MouseId) %>%
  distinct(Study, .keep_all = T) %>%
  add_count(name = "Studies") %>%
  ungroup() %>%
  mutate(CGC = HumanName %in% cgc) %>%
  mutate(COSMIC = HumanName %in% cosmic)

write.csv(export, file = "../table_app/ccgd_export.csv", row.names = F)

deployApp(appDir = "../table_app", forceUpdate = T, launch.browser = F)
```

#### build_table.R table write and app deployment snippet

<div class="notes">
- shiny snippet represents full set of code needed to render app ui. app is 200 lines total.
- lines of code replacement is orders of magnitude. possible 1:100.
- use of this script allowed for elimination of entire sql database.
</div>

```{r echo = T}
export <- homogTable %>%
  full_join(df) %>%
  group_by(MouseId) %>%
  distinct(Study, .keep_all = T) %>%
  add_count(name = "Studies") %>%
  ungroup() %>%
  mutate(CGC = HumanName %in% cgc) %>%
  mutate(COSMIC = HumanName %in% cosmic)

write.csv(export, file = "../table_app/ccgd_export.csv", row.names = F)

deployApp(appDir = "../table_app", forceUpdate = T, launch.browser = F)
```

#### build_table.R table write and app deployment snippet

@@

## Conclusions
- Stated requirements set out by this project have been met.
- A new, more modern and simplified framework has been implemented across the full application stack.
- All original functionality has been sustained.
- Design and engineering considerations were given to ease future maintenance.

## Future Directions
- Continuous integration/continuous deployment
- Anticipating future upgrades and framework evolution

@@

@@

>>>>

> - Cancer bio is complicated
> - Driver genes present primary targets
> - Screening methods...

@@

---

@@

> - Create a web tool
> - Form homology relationships
> - Curate disease classification
> - Offer connections to external tools

<div class="notes">
- Create an interactive web tool to quickly refine the stated data set by a variety of parameters.
- Draw from universal resources to define homology relationships for candidate genes between animal models.
- Curate disease classification and meta-analysis through complete CIS listing and relative rank and strength scores.
- Offer easy access to external tools from filtered results for deeper biological context.

---

- relative rank was based either on the number of insertions in a given CIS or the P-value associated with the given CIS
- The relative rank was calculated by ranking all CISs in a given study and then assigning grades
- We assigned a letter grade of A for the top 10%, B for the top 11–25%, C for the top 26–50% and D for the bottom 50%.
</div>

@@

<div class="col">
> - Filtered results are downloadable
> - Wrapped in a simple, instructional website built with classic LAMP stack
> - Hosted by UMN-OIT
</div>

@@

> - Machine displacement in provisioning by UMN-OIT
> - Use of abandoned, proprietary widgets
> - Outsized SQL database

@@

- <span style="color: #a3be8c;">**Machine displacement in provisioning by UMN-OIT**</span>
- Use of abandoned, proprietary widgets
- <span style="color: #a3be8c;">**Outsized SQL database**</span>

@@

<div class="notes">
- rshiny format is ui + server -> deployment
- server is made up of reactive constructs of inputs, expressions, and outputs
- employs declarative programming style allowing for lazy code execution.
- define loose, high level controls to passively fill functional requirements
- improves flexibility for dev and machine - resource use reduce through modularity in code execution
- our app uses 2 state path
</div>

@@

<div class="notes">
- typically, server side processing with a table of this side would be prohibitive
</div>

@@

<div class="notes">
- codebase simplified in a few ways.
- build table process was dramatically streamlined.
- shiny snippet represents full set of code needed to render app ui. app is 200 lines total.
- lines of code replacement is orders of magnitude. possibly 1:100.
- use of this script allowed for elimination of entire sql database.
</div>

@@

## Conclusions
> - Stated requirements set out by this project have been met.
> - A new, more modern and simplified framework has been implemented across the full application stack.
> - All original functionality has been sustained.
> - Design and engineering considerations were given to ease future maintenance.

## Future Directions
> - Continuous integration/continuous deployment
> - Anticipating future upgrades and framework evolution

@@

- Tim Starr*
- Yuk Sham*
- Jinhua Wang
- National Marrow Donor Program
- UMN-OIT
- Cassandra Price
- My wife and parents

@@

# Supplemental

http://hst-ccgd-prd-web.oit.umn.edu/search.html
